Interactive Curvature Tensor Visualization on Digital Surfaces\thanks{This work has been mainly funded by \textsc{DigitalSnow} ANR-11-BS02-009, \textsc{KIDICO} ANR-2010-BLAN-0205 and \textsc{PRIMES Labex} ANR-11-LABX-0063/ ANR-11-IDEX-0007 research grants.}

### 2016-01-19 latex/article.tex
minor edits
ok

\documentclass{llncs}
\usepackage{enumerate}% http://ctan.org/pkg/enumerate
\usepackage{multirow}
\usepackage{amsmath,amssymb}
\usepackage{url}
\usepackage{overpic}
\usepackage{enumerate}
\usepackage{graphicx}        % standard LaTeX graphics tool
\usepackage{tikz}        % standard LaTeX graphics tool
\usepackage{pgfplots}

@@

We first present design principles for the GPU implementation and then
present our approaches.
%\subsection{Design principles}
The general idea is to perform an Integral Invariant computation on
GPU at each vertex of the generated triangulated mesh.
\color{blue}
Since GPU's are massively parallel, we can do all those computations in parallel 
to get a very efficient implementation. Please note that
\normalcolor
when the LoD criterion is removed, each MC vertex is exactly centered at
a surfel center. At different depth of the octree, MC vertices still
correspond to surfel centers of subsampled versions of the input
object. Hence, Integral Invariant framework defined in
Section~\ref{sec:curv-tens-estim} is consistent with the triangulated
surface obtained on GPU: triangles are used for visualization purposes
but all computations are performed on the digital object $\DSh$ for
estimators
(\ref{eq:dig-curvature-estimator-k}),
(\ref{eq:dig-curvature-estimator-H}), (\ref{eq:dig-curvature-estimator-k1})
and (\ref{eq:dig-curvature-estimator-k2}).

\color{blue}
To implement the integration on $\Ball{R}{\vx} \cap \Shape$ (Fig.~\ref{fig:approx}-$a$), we tried several approaches.
\normalcolor

@@

\paragraph{Naive Approach.}
\txtblue{A first simple solution  consists in scanning all digital
points, at a given resolution, lying inside the integration domain}
(Fig.~\ref{fig:approx}-$b$) and then \txtblue{estimating} the
geometrical moment as the sum of the geometrical moments of each elementary
cubes lying in the intersection (see
Eq.~(\ref{eq:dig-curvature-estimator-k}) to (\ref{eq:dig-curvature-estimator-k2})). 
\color{blue}
This is exactly similar to what is done on the CPU. On  the GPU, we can
exploit \emph{mipmap} textures to obtain a multi-resolution information.
\normalcolor
If the input binary object is stored in a
3D texture, GPU hardware constructs a multi-resolution pyramid (mipmap) such that 8
neighboring voxel intensities at level $l$ are averaged to define the voxel
value at level $l+1$. If the level 0 corresponds to the binary input object, at
\color{blue}
a given level $l$, a texture probe returns a voxel $(x,y,z)$, containing the fraction of $\DSh$ belonging
\normalcolor
to the cube of center $(x,y,z)$ and edge length $2^l$. As a consequence, we can
approximate the volume of $\Ball{R}{\vx}\cap\Shape$ by considering mipmap values
at a given resolution $l$ (Fig.~\ref{fig:approx}-$c,d$). In this case, errors only
occurs for cells lying inside $\Shape$ (with density 1) not entirely covered by
$\Ball{R}{\vx}$. Furthermore, the \emph{mipmap} texture can be used to design,
using a single texture probe, a fast inclusion test of a given cell $c$ at level
$l$ into the shapes: we say that $c$ is in $\Shape$ if its density (retrieved
from the texture probe at level l) is greater than $1/2$. The idea here is to
mimic a kind of adaptive Gauss digitization process.

@@

We describe here two different visualization pipeline on GPU that will
be evaluated in the next section.

\paragraph{Hierarchical Decomposition.}  
\color{blue}
Using the hierarchical nature of the mipmap texture, we could also
consider hierarchical decompositions of $\Ball{R}{\vx}$
(Fig.~\ref{fig:approx2}-$a$) 
%or $\Ball{R}{\vx}\cap\Shape$
%(Fig.~\ref{fig:approx2}-$b$). 
The idea would be to decompose the
regions into mipmap cells of different resolution in order to limit
the number of texture access (important bottleneck on GPU hardwares)
and to get better approximated quantities. 
%Note that the scenario in
%Fig.~\ref{fig:approx2}-$b$ is clearly the best approach: the
%volume is exact  (\wrt the regular grid case) since geometrical moments
%are additive quantities, and the number of texture fetches is  optimal
%at this point.
%As discussed above, best
%results in terms of texture fetch can be obtained by the approach
%presented in Fig.~\ref{fig:approx2}-$b$. 
%At each vertex, the
%computation algorithm would consist in a recursive decomposition of
%the domain such that a cell is subdivided if it does not belong to
%$\Ball{R}{\vx}\cap\Shape$. However, on the GPU %hardware,
%this approach is
%highly inefficient since the hardware optimizes the parallelism only
%when the micro-program described in the shader has a predictive
%execution flow. Hence, implementing the recursive algorithm (or its
%de-recursified version) requires a lot of branches (conditional
%\emph{if} tests) in the flow. 
We have thus considered a fast multi-resolution approximation as
Fig.~\ref{fig:approx2}-$a$: for a given radius $R$, we precompute a
hierarchical decomposition of $B_R$. Such decomposition is made of
octree cells at different resolutions. At a given MC vertex $\vx$, the
algorithm becomes simple since we just scan the ball cells and
estimate the area (or other moments) from the mipmap values associated
to this cell. Note that cells in the $B_R$ decomposition may not be
aligned with mipmap cells. However, we can use GPU %mipmap
functionalities which interpolates mipmap values if we probe at
non-discrete positions. For a given radius $R$, the expected number of
cells in $B_R$ is in $O(\log{R})$. Implementation details on the
hierarchical octree representation can be found in the supplementary
material (see Section \ref{sec:discussion}).
Still, this method is not fully exploiting the GPU parallelism. 
The hardware optimizes the parallelism only
when the micro-program described in the shader has a predictive
execution flow. Hence, implementing this recursive algorithm (or its
de-recursified version) requires a lot of branches (conditional
\emph{if} tests) in the flow. 
\normalcolor

>>>>

\documentclass{llncs}
\usepackage{enumerate}% http://ctan.org/pkg/enumerate
\usepackage{multirow}
\usepackage{amsmath,amssymb}
\usepackage{url}
\usepackage{overpic}
\usepackage{enumerate}
\usepackage{graphicx}        % standard LaTeX graphics tool
\usepackage{tikz}        % standard LaTeX graphics tool
\usepackage{pgfplots}
\usepackage[normalem]{ulem}

@@

We first present design principles for the GPU implementation and then
present our approaches.
%\subsection{Design principles}
The general idea is to perform an Integral Invariant computation on
GPU at each vertex of the generated triangulated mesh.
\color{blue}
Since GPU have massively parallel architectures, we can do all those computations in parallel 
to obtain a very efficient implementation. Please note that
\normalcolor
when the LoD criterion is removed, each MC vertex is exactly centered at
a surfel center. At different depth of the octree, MC vertices still
correspond to surfel centers of subsampled versions of the input
object. Hence, Integral Invariant framework defined in
Section~\ref{sec:curv-tens-estim} is consistent with the triangulated
surface obtained on GPU: triangles are used for visualization purposes
but all computations are performed on the digital object $\DSh$ for
estimators
(\ref{eq:dig-curvature-estimator-k}),
(\ref{eq:dig-curvature-estimator-H}), (\ref{eq:dig-curvature-estimator-k1})
and (\ref{eq:dig-curvature-estimator-k2}).

\txtblue{To implement the integration on $\Ball{R}{\vx} \cap \Shape$
  (Fig.~\ref{fig:approx}-$a$), several strategies have been evaluated.}

@@

\paragraph{Naive Approach.}
\txtblue{A first simple solution  consists in scanning all digital
points, at a given resolution, lying inside the integration domain}
(Fig.~\ref{fig:approx}-$b$) and then \txtblue{estimating} the
geometrical moment as the sum of the geometrical moments of each elementary
cubes lying in the intersection (see
Eq.~(\ref{eq:dig-curvature-estimator-k}) to (\ref{eq:dig-curvature-estimator-k2})). 
\color{blue}
This is exactly similar to what is done on the CPU. On  the GPU, we can
exploit \emph{mipmap} textures to obtain  multi-resolution information.
\normalcolor
If the input binary object is stored in a
3D texture, GPU hardware constructs a multi-resolution pyramid (mipmap) such that 8
neighboring voxel intensities at level $l$ are averaged to define the voxel
value at level $l+1$. If the level 0 corresponds to the binary input object, at
\color{blue}
a given level $l$, a texture probe at  a point $(x,y,z)$ returns the fraction of $\DSh$ belonging
\normalcolor
to the cube of center $(x,y,z)$ and edge length $2^l$. As a consequence, we can
approximate the volume of $\Ball{R}{\vx}\cap\Shape$ by considering mipmap values
at a given resolution $l$ (Fig.~\ref{fig:approx}-$c,d$). In this case, errors only
occurs for cells lying inside $\Shape$ (with density 1) not entirely covered by
$\Ball{R}{\vx}$. Furthermore, the \emph{mipmap} texture can be used to design,
using a single texture probe, a fast inclusion test of a given cell $c$ at level
$l$ into the shapes: we say that $c$ is in $\Shape$ if its density (retrieved
from the texture probe at level l) is greater than $1/2$. The idea here is to
mimic a kind of adaptive Gauss digitization process.

@@

\txtblue{\sout{We describe here two different visualization pipeline on GPU that will
be evaluated in the next section.}}

\paragraph{Hierarchical Decomposition.}  
\color{blue}
Using the hierarchical nature of the mipmap texture, we could also
consider hierarchical decompositions of $\Ball{R}{\vx}$
(Fig.~\ref{fig:approx2}-$a$) 
%or $\Ball{R}{\vx}\cap\Shape$
%(Fig.~\ref{fig:approx2}-$b$). 
The idea is to decompose the
ball into mipmap cells of different resolution in order to limit
the number of texture access (important bottleneck on GPU hardwares)
and to get better approximated quantities. 
%Note that the scenario in
%Fig.~\ref{fig:approx2}-$b$ is clearly the best approach: the
%volume is exact  (\wrt the regular grid case) since geometrical moments
%are additive quantities, and the number of texture fetches is  optimal
%at this point.
%As discussed above, best
%results in terms of texture fetch can be obtained by the approach
%presented in Fig.~\ref{fig:approx2}-$b$. 
%At each vertex, the
%computation algorithm would consist in a recursive decomposition of
%the domain such that a cell is subdivided if it does not belong to
%$\Ball{R}{\vx}\cap\Shape$.
\txtred{On the GPU, computing a hierarchical decomposition of
  $B_R(\vx)$ at each point $\vx$ is highly inefficient since the
  hardware optimizes the parallelism only when the micro-program
  described in the shader has a predictive execution flow. Hence,
  implementing the recursive algorithm (or its de-recursified version)
  requires a lot of branches (conditional \emph{if} tests) in the
  flow.}  We have thus considered a fast multi-resolution
approximation as illustrated in Fig.~\ref{fig:approx2}-$a$: for a
given radius $R$, we precompute a hierarchical decomposition of
$B_R$. Such decomposition is made of octree cells at different
resolutions. At a given MC vertex $\vx$, the algorithm becomes simple
since we just scan the ball cells and estimate the area (or other
moments) from the mipmap values associated to this cell. Note that
these \txtred{precomputed} cells in the $B_R$ decomposition may not be aligned
with mipmap cells. However, we can use GPU %mipmap functionalities
which interpolates mipmap values if we probe at non-discrete
positions. For a given radius $R$, the expected number of cells in
$B_R$ is in $O(\log{R})$. Implementation details on the hierarchical
octree representation can be found in the supplementary material (see
Section \ref{sec:discussion}). \sout{ Still, this method is not fully
exploiting the GPU parallelism.  The hardware optimizes the
parallelism only when the micro-program described in the shader has a
predictive execution flow. Hence, implementing this recursive
algorithm (or its de-recursified version) requires a lot of branches
(conditional \emph{if} tests) in the flow.}\todo{la discussion vient apres}  \normalcolor
